Sparse Function-space Representation \\ of Neural Networks

### 2023-05-23 paper/main.tex
App. C

\section{\our for model-based reinforcement learning}
\label{app:rl}
In this appendix, we provide more details of the reinforcement learning experiments, including
how we used \our in this setting.

\subsection{Background}
We consider environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expectation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1)$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics is learned \(f_{\mathbf{w}^{*}} \approx \transitionFn\) and then used by a planning algorithm.
A simple approach is to use the learned dynamic model $f_{\mathbf{w}^{*}}$ and maximise the objective in \cref{eq-model-free-objective},
\begin{align} \label{eq-greedy}
  \policy^{\text{Greedy}} &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi).
\end{align}
However, we can leverage the method in \cref{sec:methods} to obtain a function-space posterior over the learned dynamics $q_{\mathbf{u}}(\transitionFn \mid \dataset)$,
where $\mathcal{D}$ represents the state transition data set \(\mathcal{D} = \{(s_{i},a_{i}), s_{i+1}\}_{i=0}^{N}\).
Importantly, the uncertainty represented by this posterior distribution can be used to balance the exploration-exploitation trade-off,
using approaches such as posterior sampling \cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013},
\begin{align} \label{eq-posterior-sampling}
  \policy^{\text{PS}} &= \arg \max_{\pi \in \Pi} J(\tilde{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}({\transitionFn} \mid \dataset),
\end{align}
where a function $\tilde{\transitionFn}$ is sampled from the (approximate) posterior $q_{\mathbf{u}}({\transitionFn} \mid \dataset)$ and used to find a policy as
in \cref{eq-greedy}.
Intuitively, this strategy will explore where the model has high uncertainty, which in turn will reduce the model's uncertainty as data is collected and used to
train the model.
This strategy should perform well in environments which are hard to explore, for example, those
with sparse rewards.

@@

\subsection{Algorithm}
In practice, it is common to approximate the infinite sum in \cref{eq-posterior-sampling}
by planning over a finite
horizon and bootstrapping with a learned value function.
Following this approach, our planning algorithim is given by,
\begin{align} \label{eq-fast-update-mpc}
  \policy^{\text{PS}}(\state) = \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
\E \bigg[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state  \bigg] + Q_{\theta}(\state_{\Horizon}, \action_{H})
\quad \text{s.t. } \state_{t+1} &= \tilde{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t}
\end{align}
where $\tilde{\transitionFn} \sim q_{\mathbf{u}}(\transitionFn \mid \dataset)$ is a sample from the
dynamics posterior and
${Q_{\theta}(\state, \action) \approx \E \big[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state, \action_{0}=\action  \big]}$
is a learned value function with parameters $\theta$.
We use deep deterministic policy gradient (DDPG) \cite{lillicrapContinuousControlDeep2016} to learn the action value function $Q_{\theta}$.
Note that we also learn a policy but its sole purpose is for learning the value function.

@@

\subsection{Environment}
We consider the cart pole swing up task in MuJoCo \cite{todorov2012mujoco}.
However, we make exploration difficult by implementing a sparse reward function which returns $0$ unless the reward is over a threshold value. That is, our reward function is given by,
$$\hat{r}(\state_{t}, \action_{t}) =
\begin{cases}
    r(\state_{t}, \action_{t}),& \text{if } r(\state_{t}, \action_{t})\geq 0.6\\
    0,              & \text{otherwise}
\end{cases}
$$
In all experiments we collected an initial data set using a random policy for one episode and
we set action repeat as two.

\subsection{Experiment Configuration}
This section details how we configured and ran our reinforcement learning experiments.

@@

\textbf{Ensemble}
The ensemble experiments used $5$ members where each member was an MLP with two hidden layers of width $128$ and Tanh activations, i.e. the same architecture as the other experiments.
At each episode, the ensemble was trained for $10000$ iterations using the Adam optimiser.
We used a learning rate $0.0001$ and a batch size of $64$.
At each time step in the planning horizon we used the following procedure to sample from the ensemble.
First, the mean and variance of the ensemble members predictions were calculated.
Next, we sampled from the resulting Gaussian.
\cref{fig:rl} shows that the ensemble required more environment interaction to solve the task.
This is indicated by the green line not converging within the $50$ episodes shown in the figure.

>>>>

\section{\our for model-based reinforcement learning}
\label{app:rl}
%
As a final experiment in the main paper in \cref{sec:rl-exp}, we provided a demonstration of combining our \our approach with model-based reinforcement learning. As such, this is a direct extension of the uncertainty quantification and representation capabilities already demonstrated in the supervised (and continual) learning examples in the main paper. However, the RL setting requires slightly more background to it and thus \cref{app:rl} has been separated out to a separate section in the appendix. In this appendix, we provide details of the reinforcement learning experiments, including how we used \our in this setting.

\subsection{Background}
We consider environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d.\ transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expectation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1)$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics is learned \(f_{\mathbf{w}^{*}} \approx \transitionFn\) and then used by a planning algorithm.
A simple approach is to use the learned dynamic model $f_{\mathbf{w}^{*}}$ and maximise the objective in \cref{eq-model-free-objective},
\begin{align} \label{eq-greedy}
  \policy^{\text{Greedy}} &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi).
\end{align}
However, we can leverage the method in \cref{sec:methods} to obtain a function-space posterior over the learned dynamics $q_{\mathbf{u}}(\transitionFn \mid \dataset)$,
where $\mathcal{D}$ represents the state transition data set \(\mathcal{D} = \{(s_{i},a_{i}), s_{i+1}\}_{i=0}^{N}\).
Importantly, the uncertainty represented by this posterior distribution can be used to balance the exploration-exploitation trade-off,
using approaches such as posterior sampling \cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013},
\begin{align} \label{eq-posterior-sampling}
  \policy^{\text{PS}} &= \arg \max_{\pi \in \Pi} J(\tilde{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}({\transitionFn} \mid \dataset),
\end{align}
where a function $\tilde{\transitionFn}$ is sampled from the (approximate) posterior $q_{\mathbf{u}}({\transitionFn} \mid \dataset)$ and used to find a policy as
in \cref{eq-greedy}.
Intuitively, this strategy will explore where the model has high uncertainty, which in turn will reduce the model's uncertainty as data is collected and used to
train the model.
This strategy should perform well in environments which are hard to explore, for example, those
with sparse rewards.

@@

\subsection{Algorithm}
In practice, it is common to approximate the infinite sum in \cref{eq-posterior-sampling}
by planning over a finite
horizon and bootstrapping with a learned value function.
Following this approach, our planning algorithim is given by,
\begin{equation} \label{eq-fast-update-mpc}
  \policy^{\text{PS}}(\state) = \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
\E \bigg[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state  \bigg] + Q_{\theta}(\state_{\Horizon}, \action_{H}),
%\quad \text{s.t. } \state_{t+1} &= \tilde{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t}
\end{equation}
such that $\state_{t+1} = \tilde{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t}$, 
where $\tilde{\transitionFn} \sim q_{\mathbf{u}}(\transitionFn \mid \dataset)$ is a sample from the dynamics posterior and
${Q_{\theta}(\state, \action) \approx \E \big[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state, \action_{0}=\action  \big]}$
is a learned value function with parameters $\theta$.
We use deep deterministic policy gradient (DDPG, \cite{lillicrapContinuousControlDeep2016}) to learn the action value function $Q_{\theta}$.
Note that we also learn a policy but its sole purpose is for learning the value function.

@@

\subsection{Environment}
We consider the cartpole swingup task in MuJoCo~\cite{todorov2012mujoco}.
However, we make exploration difficult by implementing a sparse reward function which returns $0$ unless the reward is over a threshold value. That is, our reward function is given by,
$$\hat{r}(\state_{t}, \action_{t}) =
\begin{cases}
    r(\state_{t}, \action_{t}),& \text{if } r(\state_{t}, \action_{t})\geq 0.6\\
    0,              & \text{otherwise}
\end{cases}
$$
In all experiments we collected an initial data set using a random policy for one episode and we set action repeat as two.

\subsection{Experiment configuration}
%
This section details how we configured and ran our reinforcement learning experiments.

@@

\textbf{Ensemble}
The ensemble experiments used $5$ members where each member was an MLP with two hidden layers of width $128$ and Tanh activations, \ie\ the same architecture as the other experiments.
At each episode, the ensemble was trained for $10000$ iterations using the Adam optimiser.
We used a learning rate $0.0001$ and a batch size of $64$.
At each time step in the planning horizon we used the following procedure to sample from the ensemble.
First, the mean and variance of the ensemble members predictions were calculated.
Next, we sampled from the resulting Gaussian.
\cref{fig:rl} shows that the ensemble required more environment interaction to solve the task.
This is indicated by the green line not converging within the $50$ episodes shown in the figure.
